---
title: "Piece 1"
author: "Benjamin Egan"
date: "2025-01-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Ok so I did this a year ago when I was trying things and now I no longer remember which packages I need. So I left them all in. I hate it too...
library(haven)
library(lme4)
library(lmerTest)
library(psych)
library(emmeans)
library(ltm)
library(Matrix)
library(sjPlot)
library(reshape2)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(survMisc)
library(hrbrthemes)
library(geomtextpath)

setwd("/Users/egan/Desktop/Gradschool/My Research")
data <- read_spss("Current Dataset 2-28.sav")
view(data)

```



## Family Decision Making


### What is Family Decision Making

Adolescents were presented with a variety o topics that families with adolescents often have to make decisions about, and chose the answer that best reflected who made most of the decisions on each topic. The scale ranged from 1 (My parents decide this without discussing it with me) to 5 (I decide this without discussing it with my parents), with 3 representing a joint decision (My parents and I make the decision together).



```{r code to get it up to speed, echo = false}
data$Country <- as.factor(data$IDN)
Country <- factor(data$Country) #country into categorical variable
data$AgeCM <- data$Age - mean(data$Age, na.rm=TRUE) #mean center age
data$FDMCM <- data$FDM - mean(data$FDM, na.rm=TRUE) #mean center age



#change values for males from 1 to 0 
data$ChGender <- replace(data$ChGender, data$ChGender < 2, 0) 
#change values for females from 2 to 1 
data$ChGender <- replace(data$ChGender, data$ChGender > 0, 1) 


#renaming Countries
levels(data$Country)[levels(data$Country)=='1'] <- 'China-Jinan'
levels(data$Country)[levels(data$Country)=='2'] <- 'China-Shanghai'
levels(data$Country)[levels(data$Country)=='3'] <- 'Italy-Naples'
levels(data$Country)[levels(data$Country)=='4'] <- 'Italy-Rome'
levels(data$Country)[levels(data$Country)=='5'] <- 'Kenya'
levels(data$Country)[levels(data$Country)=='6'] <- 'Philippines'
levels(data$Country)[levels(data$Country)=='7'] <- 'Thailand'
levels(data$Country)[levels(data$Country)=='8'] <- 'Sweden'
levels(data$Country)[levels(data$Country)=='9'] <- 'USA-African American'
levels(data$Country)[levels(data$Country)=='10'] <- 'USA-European American'
levels(data$Country)[levels(data$Country)=='11'] <- 'USA-Hispanic'
levels(data$Country)[levels(data$Country)=='12'] <- 'Colombia'
levels(data$Country)[levels(data$Country)=='13'] <- 'Jordan'
levels(data$Country)[levels(data$Country)=='16'] <- 'India'
levels(data$Country)[levels(data$Country)=='17'] <- 'China-Chongqing'
levels(data$Country)
data$Country <- relevel(data$Country, ref = 'USA-Hispanic')

```


