---
title: "Piece 1"
author: "Benjamin Egan"
date: "2025-01-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Ok so I did this a year ago when I was trying things and now I no longer remember which packages I need. So I left them all in. I hate it too...
library(haven)
library(lme4)
library(lmerTest)
library(psych)
library(emmeans)
library(ltm)
library(Matrix)
library(sjPlot)
library(reshape2)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(survMisc)
library(hrbrthemes)
library(geomtextpath)

data <- read_spss("FDM Dataset for Data Science.sav")

```



## Tragectories and Variations of Family Decision Making

In this piece, I will demonstrate my graphing abilities.

### What is Family Decision Making

Adolescents were presented with a variety o topics that families with adolescents often have to make decisions about, and chose the answer that best reflected who made most of the decisions on each topic. The scale ranged from 1 (My parents decide this without discussing it with me) to 5 (I decide this without discussing it with my parents), with 3 representing a joint decision (My parents and I make the decision together).


```{r seting up FDM, echo = FALSE}

data$Country <- as.factor(data$IDN)
Country <- factor(data$Country) #country into categorical variable
data$AgeCM <- data$Age - mean(data$Age, na.rm=TRUE) #mean center age
data$FDMCM <- data$FDM - mean(data$FDM, na.rm=TRUE) #mean center FDM

#change values for males from 1 to 0: data$ChGender <- replace(data$ChGender, data$ChGender < 2, 0) 
#change values for females from 2 to 1: data$ChGender <- replace(data$ChGender, data$ChGender > 0, 1) 

```

```{r Renaming Countries, echo=FALSE}

#renaming Countries
levels(data$Country)[levels(data$Country)=='1'] <- 'China-Jinan'
levels(data$Country)[levels(data$Country)=='2'] <- 'China-Shanghai'
levels(data$Country)[levels(data$Country)=='3'] <- 'Italy-Naples'
levels(data$Country)[levels(data$Country)=='4'] <- 'Italy-Rome'
levels(data$Country)[levels(data$Country)=='5'] <- 'Kenya'
levels(data$Country)[levels(data$Country)=='6'] <- 'Philippines'
levels(data$Country)[levels(data$Country)=='7'] <- 'Thailand'
levels(data$Country)[levels(data$Country)=='8'] <- 'Sweden'
levels(data$Country)[levels(data$Country)=='9'] <- 'USA-African American'
levels(data$Country)[levels(data$Country)=='10'] <- 'USA-European American'
levels(data$Country)[levels(data$Country)=='11'] <- 'USA-Hispanic'
levels(data$Country)[levels(data$Country)=='12'] <- 'Colombia'
levels(data$Country)[levels(data$Country)=='13'] <- 'Jordan'
levels(data$Country)[levels(data$Country)=='16'] <- 'India'
levels(data$Country)[levels(data$Country)=='17'] <- 'China-Chongqing'
data$Country <- relevel(data$Country, ref = 'USA-Hispanic')

```
### Data Being Used 

Several different variables were used for the following analyses. Alongside Family Decision Making (FDM), Age (AgeCM), Socio-economic Status (SES), Gender (ChGender), and Country of origin were used to help identify potential trends in the data. Below are the countries utilized in this dataset. For now, I will only be working with people from the United States. 

```{r showing data}
levels(data$Country)

```
### US Dataset Creation


```{r US data setup}

USdata <- subset(data,IDN ==9 | IDN == 10 | IDN == 11)
USdata$Country <- as.factor(USdata$IDN)
Country <- factor(data$Country) #country into categorical variable
levels(USdata$Country)[levels(USdata$Country)=='9'] <- 'USA-African American'
levels(USdata$Country)[levels(USdata$Country)=='10'] <- 'USA-European American'
levels(USdata$Country)[levels(USdata$Country)=='11'] <- 'USA-Hispanic'
USdata$Country <- relevel(USdata$Country, ref = 'USA-European American')
USdata$AgeCM <- USdata$Age - mean(USdata$Age, na.rm=TRUE) #mean center age
USdata$FDMCM <- USdata$FDM - mean(USdata$FDM, na.rm=TRUE) #mean center age
data$Age_Exact <- round(data$Age, digits = 0)

```

## Plotting Family Decision Making in the United States


```{r multilevel model, echo = FALSE}
#TEXT FOR IF I INCLUDE MY MODEL:   My first step was to create a multi-level model. The model below uses Family Decision Making as the outcome. The inputs were gender, country, and an interaction between age and country.

#model I used for FDM. Looking to incorporate this somehow? Not sure how.
model.FDM_USA <- lmer(FDM ~ SES + ChGender + AgeCM + Country +   + Country*AgeCM + (1|ID), data = USdata)
```




I can now create a line graph to plot the multilevel model.

```{r plot model}




USdata %>%
ggplot(
  aes(
    x=Age,
    y = FDM,
    group = Country, 
    color = Country,
    )) + 
  geom_smooth(se = FALSE) +
  theme_bw() +
  scale_color_manual(values=c("#0000ff", "#ff0000","#008000")) +
  labs(
    x = "Age", 
    y = "Family Decision Making", 
    color = "Ethnicity",
    title = "USA Age and Ethnicity Effects on Family Decision Making"
    )

```
